---
title: "p8105_hw6_kz2334"
author: "Kangkang Zhang"
output: github_document
---

Load required packages.

```{r set up, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(viridis)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


##Problem 1

Import the data. 
```{r 1.1 import, message = FALSE} 
homicide_data = read_csv("./data/homicide-data.csv")
```

Create a city_state variable, and a binary variable indicating whether the homicide is solved. 

```{r 1.2}
homicide_data = homicide_data %>% 
  mutate(city_state = str_c(city, ", ", state),
         status = ifelse(disposition %in% c("Closed without arrest", "Open/No arrest"), 0, 1))
```

Omit cities Dallas, TX; Phoenix, AZ; Kansas City, MO and Tulsa, AL. Modifiy victim_race to have categories white and non-white, with white as the reference category. Convert victim_age to numeric.

```{r 1.3, warning = FALSE}
homicide_md = homicide_data %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>% 
  mutate(victim_race = ifelse(victim_race == "White", "white", "non-white"),
         victim_race = as.factor(victim_race), 
         victim_race = relevel(victim_race, ref = "white"),
          victim_age = as.numeric(victim_age))
  
```

Run logistic regression for the city of Baltimore, MD. Find out the CI and estimate of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r 1.4, message = FALSE}
balt_logistic = homicide_md %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(status ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

balt_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
    conf_low_or = exp(confint(balt_logistic, level = 0.95)[,1]),
    conf_high_or = exp(confint(balt_logistic, level = 0.95)[,2])) %>% 
  filter(term == "victim_racenon-white") %>% 
  select(OR, starts_with("conf")) %>% 
  knitr::kable(digit = 3)
```

We can see that adjusted OR for solving homicides comparing non-white victims to white victims is 0.441. 95% CI is (0.312, 0.620). The OR solving homicides for non-white victims is 0.441 times of that for white victims, which means non-white victims are less likely to be with solving homicides.

